Towards a Sound and Complete Library for Modal Logics Reasoning in Coq

### 2022-06-20 paper/paper.tex

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  This work describes an encoding of modal logics using the Coq proof assistant.
  Our formalization differs from previous attempts by providing a uniform
  representation of several systems for modal logics using Coq's rich type
  structure. We illustrate the usefulness of our library in a formalization of
  L\"ob's theorem which closely follows a classical proof of this result.
\end{abstract}

@@

The next table summarizes the different systems of modal logics considered by
previous works.

\vspace{0.5cm} {
  \begin{table}[H]
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
      \hline
      \multirow{2}{*}{Authors} & \multicolumn{9}{c|}{Systems considered} \\ \cline{2-10} 
                               & K & D & B & T & K4 & K5 & S4 & S5 & GL\\ \hline
      Wind et. al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X & \\ \hline
      Doczkal et. al.~\cite{saarland}   & X &   &   & X & X  &    &    &  &  \\ \hline
      Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
      This work         & X & X & X & X & X  &  X &  X & X  & X \\ \hline
    \end{tabular}
    \centering
    \caption{Comparision with different Coq modal logics developements.}
    \label{tab:comparacao_trabalhos_rel}
  \end{table}
}

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
  This work describes an encoding of modal logics using the Coq proof assistant.
  Our formalization differs from previous attempts by providing a uniform
  representation of several systems for modal logic using Coq's rich type
  structure. We illustrate the usefulness of our library in a formalization of
  L\"ob's theorem which closely follows a classical proof of this result.
\end{abstract}

@@

The next table summarizes the formalized modal logics systems by our library
and related works.
\vspace{0.5cm} {
  \begin{table}[H]
    \begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
      \hline
      \multirow{2}{*}{Authors} & \multicolumn{9}{c|}{Systems considered} \\ \cline{2-10} 
                               & K & D & B & T & K4 & K5 & S4 & S5 & GL\\ \hline
      Wind et. al.~\cite{dewind}     & X &   &   & X & X  & X  & X  &  X & \\ \hline
      Doczkal et. al.~\cite{saarland}   & X &   &   & X & X  &    &    &  &  \\ \hline
      Benzmüller et. al.~\cite{Chris_Bruno}& X &   & X & X &  X &    &    & X &  \\ \hline
      This work         & X & X & X & X & X  &  X &  X & X  & X \\ \hline
    \end{tabular}
    \centering
    \caption{Comparision with different Coq modal logics developements.}
    \label{tab:comparacao_trabalhos_rel}
  \end{table}
}
All developments considered systems K, T and K4 and the unique formalization that
includes system GL is the library described in this work.

\paragraph{Applications of modal logics in semantics} Another application of
modal logics is its use for reasoning about programming languages semantics.
Dynamic logic was introduced by Pratt in late 70's as modal logic for reasoning
about programs~\cite{Pratt76,Pratt80}. The key idea on dynamic logic is to express the
Hoare triples by a modality meaning ``after a program $p$ executes, formula
$\varphi$ will be true''. The programming languages considered in Pratt's
original work are composed by an assignment and regular language operators
(union, concatenation and Kleene star)~\cite{Pratt76}.

An application of modal logic IS5 as a type system for a distributed
programming language was proposed in~\cite{Murphy04}. 
In that work, modalities are used to represent types for mobile code
and remote addresses. Murphy et. al defined the language by interpreting
the possible worlds of the intutionistic modal logic IS5 to be nodes on a
network and the modalities $\Box$ and $\Diamond$ to reflect mobility and
locality, respectively. Authors develop an operational semantics and show
that their formalization is type-safe.

The application of a multimodal separation logic for the semantics of low-level
was the subject of Dockins et. al. work~\cite{Dockins08}. In that work, authors
combine several modal operators that do not need to satisfy the same frame
conditions and seperation logics for reasoning about non-trivial subsets of
low-level programming languages. Authors developed their formalization through a
shallow embedding in Coq proof assistant and provide two case studies: one is
the verification of a simple low-level virtual machine, called the list machine
benchmark and the other is the subset of Sequential Cminor.

A model for recursive, impredicatively quantified types with mutable references
was studied by Appel et. al.~\cite{Appel07}. His interest in formalizing such
advanced typing concepts was in the context of typed assembly languages.
Authors stabilished a semantic type soundness proof by defining a Kripke-style 
semantics of the G\"odel-L\"ob logic of provability. All their definitions were
mechanized through a shallow embedding in Coq.

The embedding of modal linear logic in Coq was the subject of
Sadrzadeh's work~\cite{Soton03}. Like our work, Sadrzadeh's uses a
deep-embedding approach and assumes that all modalities are reflexive
and transitive. Using her embedding, Sadrzadeh shows some proofs of classical
logic puzzles.
